---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LidarIndexR

<!-- badges: start -->
<!-- badges: end -->

The goal of LidarIndexR is to provide functions to build index files for lidar projects. The functions
in the package have been tested using FTP and HTTPS servers. However, I can't guarantee that the directory
functions work with all HTTP(S) servers due to differences in the HTML generated to show directory contents.

## Installation

You can install the released version of LidarIndexR from [CRAN](https://CRAN.R-project.org) with:

``` r
# install.packages("LidarIndexR")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bmcgaughey1/LidarIndexR")
```
## Example #1: Create a tile index from files on a remote ftp server

This example creates a tile index for a single lidar project stored in the USGS rockyweb server. In this case,
you have to know the details for the project including the folder name for the project and the projection
information for the data.

```{r example1}
library(LidarIndexR)
library(sf)
library(ggplot2)
library(viridis)

## Basic example to build an index
URL <- "https://rockyweb.usgs.gov/vdelivery/Datasets/Staged/Elevation/LPC/Projects/AK_BrooksCamp_2012/"
pointFolder <- "laz"
outputFile <- "~/AK_BrooksCamp_2012.gpkg"

# projection info...for most projects, this can be pulled from the WESM project index
pointCRS <- 26905

# create index
BuildIndexFromPoints(URL, pointFolder, outputFile, projString = pointCRS, 
                     appendInfo = data.frame("Project" = "AK_BrooksCamp_2012"))

# read the index and display
index <- st_read(outputFile)

ggplot(index) +
  ggtitle("AK_BrooksCamp_2012 lidar tiles") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_sf(aes(fill = PointCount), show.legend = TRUE) +
  scale_fill_viridis()
 
```

## Example #2: Create a tile index using a project from the USGS WESM project index

This example creates an index for the same lidar project as Example #1. However, this
example uses the USGS project index to populate the new tile index with information
regarding the acquisition and to provide projection information for the project.

You will need to change the line that reads the index to reflect the location of the WESM index
on your computer.

The example requires a local copy of the USGS project index from
[here](https://rockyweb.usgs.gov/vdelivery/Datasets/Staged/Elevation/metadata/WESM.gpkg).
The commented code in the example downloads the WESM.gpkg index.

**Note:** The call to \code{BuildIndexFromUSGSProjectIndexItem} can take quite a bit of
time to execute. It touches every point file on the server to read bounding box information
for the tiles. For projects with thousands of tiles, it can take hours to build the index.

```{r example2}
library(LidarIndexR)
library(sf)
library(utils)
library(ggplot2)
library(viridis)

# download the USGS WESM index
# utils::download.file("https://rockyweb.usgs.gov/vdelivery/Datasets/Staged/Elevation/metadata/WESM.gpkg",
#                      "WESM.gpkg", 
#                      method = "libcurl", 
#                      mode = "wb")

# read the FESM index from a local file
index <- st_read("G:/R_Stuff/EntwineIndex/WESM.gpkg")
pointFolder <- "laz"
outputFile <- "~/AK_BrooksCamp_2012_Proj.gpkg"

item <- index[which(index$workunit == "AK_BROOKSCAMP_2012"), ]

# create index
BuildIndexFromUSGSProjectIndexItem(item, pointFolder, outputFile)

# read the index and display
tindex <- st_read(outputFile)

ggplot(tindex) +
  ggtitle("AK_BrooksCamp_2012 lidar tiles") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_sf(aes(fill = PointCount), show.legend = TRUE) +
  scale_fill_viridis()
 
```

## Example #3: Create an overall polygon for a project from the tile index

This example builds on Example #2 to create a polygon showing the area covered by the point tiles.

```{r example3}
# build on the last example to generate a project polygon

tproject <- BuildProjectPolygonFromIndex(
  outputFile,
  item$workunit,
  appendInfo = item[, c("workunit_id", "collect_start", "collect_end", "p_method", "horiz_crs")],
  quiet = FALSE
)

# display the project polygon
ggplot(tproject) +
  ggtitle(paste0(item$workunit, " project area")) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_sf(aes(fill = PID), show.legend = FALSE)
```

```{r eval=FALSE, include=FALSE}
# You will need to change the line that reads the index to reflect the location of the WESM index
# on your computer.

library(LidarIndexR)
library(sf)

# test code...not to be run as part of README

index <- st_read("G:/R_Stuff/PlotClipping/WESM.gpkg")
outputFile <- "~/AK_BrooksCamp_2012.gpkg"
appendCols <- c("opr_category", "lpc_pub_date")

item <- index[which(index$workunit == "AK_BROOKSCAMP_2012"), ]

# read the index
tindex <- st_read(outputFile)

info <- data.frame(item[, appendCols])

# drop geometry
info$geometry <- NULL

# duplicate rows
info <- do.call("rbind", replicate(nrow(tindex), info, simplify = FALSE))

# cbind to tile data frame
tindex <- cbind(tindex, info)

```
